---
title: "Project 1 - Travel Insurance Data"
author: "Kevin Lopez, Clarisse Sarah, and Mayank Thacker" 
output: word_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Work break down:

### Kevin

-   Create shared Rmd file and combined code from team members. [GitHub Repository](https://github.com/KevinDLopez/CECS450-Project1)
-   Created Report using Rmd ( Not much work, Started Rmd to be well formated for word).
-   Created Graph 4 and Graph 5 6.

### Clarisse

-   Created Graph 3.

-   Worked on [PPT](https://docs.google.com/presentation/d/1Rc63oOnZWbTFga_m7iyCOpVMNwukj7gsjbV8HFEq7xQ/edit?pli=1#slide=id.g25f6af9dd6_0_0), and report.

### Mayank

-   Created Graph 1, 2.
-   Created Worked on [PPT](https://docs.google.com/presentation/d/1Rc63oOnZWbTFga_m7iyCOpVMNwukj7gsjbV8HFEq7xQ/edit?pli=1#slide=id.g25f6af9dd6_0_0)

# Explain the dataset

The travel insurance dataset provides valuable information about travelers, offering insights into their travel insurance purchasing decisions. In addition, the dataset includes details such as age, income, employment status, and more to highlight factors influencing insurance choices. With 10 columns and 1986 rows, the data offers robust information for conducting insightful analyses and patterns within the dataset. 

### The columns are:

**Age**- Age Of The Customer

**Employment Type**- The Sector In Which Customer Is Employed

**GraduateOrNot**- Whether The Customer Is College Graduate Or Not

**AnnualIncome**- The Yearly Income Of The Customer In Indian Rupees[Rounded To Nearest 50 Thousand Rupees]

**FamilyMembers**- Number Of Members In Customer's Family

**ChronicDisease**- Whether The Customer Suffers From Any Major Disease Or Conditions Like Diabetes/High BP or Asthama,etc.

**FrequentFlyer**- Derived Data Based On Customer's History Of Booking Air Tickets On Atleast 4 Different Instances In The Last 2 Years[2017-2019].

**EverTravelledAbroad**- Has The Customer Ever Travelled To A Foreign Country[Not Necessarily Using The Company's Services]

**TravelInsurance**- Did The Customer Buy Travel Insurance Package During Introductory Offering Held In The Year 2019.

# Load the Dataset

```{r include=TRUE}
data = read.csv("./TravelInsurancePrediction.csv")
#knitr::kable(head(data), caption = "Table 1: First 6 Rows of the Dataset")
# colnames(data)
library(flextable)
library(ggplot2)
library(reshape2)

display_table = function(data){
  # Function to display a nice looking table using flextable. 
  ## It applies a minimal theme (Shows the vertical and horizontal lines of the table).
  ## https://ardata-fr.github.io/flextable-book/
  my_table <- head(data)  
  table = flextable(my_table)
  table = fontsize(table, size = 7, part = "all")
  table = align(table, align = "center", part = "all")
  table = theme_box(table)  # Apply a predefined theme ( Simple table )
  table = align(table, align = "center", part = "all")
  table = bold(table, part = "header")  # Bold text in header
  table = width(table, width = .7)
  table = width(table, j = 1, width = .2)  # Adjust width of the first column
  return(table)  # Display the flextable
}

display_table(data)


```

## Reshaping data

Data is already in long format. Data for income is given in Indian Rupees. We will create a new column to convert it to USD. Column name will be AnnualIncomeUSD

```{r include=TRUE}
RUPEE_TO_USD = 0.012
# add a new column for income in USD
data$AnnualIncomeUSD <- data$AnnualIncome * RUPEE_TO_USD
table = display_table(data)

# hihglight the last column foreground
table = color(table, color = "orange", part = "all", j = 11)
table

```


# Graph 1 - Frequency of Age Distribution
- The histogram shows the frequency of each individual age in the dataset.  It provides insights into the prevalence of different age groups within the dataset.
- 28 years old has the highest frequency
```{r}

ggplot(data, aes(x = factor(Age))) +
  geom_bar(fill = "lightcoral", color = "black") + # create histogram based on Age. 
  labs(title = "Frequency of Each Age", x = "Age", y = "Frequency") # add title and labels for x,y axis
  
```


# Graph 2 - Annual Income Density
- To visualize the probability density of a Annual Income 
- X-axis: "Annual Income", Y-axis: "Density". 

```{r}
ggplot(data, aes(x = AnnualIncome)) + # Set the x-axis to AnnualIncome
  geom_density(fill = "lightcoral", color = "black") + # Add a density layer with lightcoral fill and black border
  labs(title = "Density Plot of Annual Income", x = "Annual Income", y = "Density") # Add title and axis labels
  
```

# Graph 3 - Bar Plot of Frequent Flyer by age

- The graph highlights the distribution of flyer status across different age groups. 
- It’s evident that there is a higher number of non-frequent flyers than that of frequent flyers between the ages of 25 to 35. 
- We can see that there are fluctuations in the frequency of both frequent and non-frequent flyers. 
- In particular, there is a decline in the number of frequent flyers between ages 25 to 27 and 28 to 30 and a steady increase in non-frequent flyers. In comparison, between ages 32 to 24, there is an increase in the number of frequent flyers and non-frequent flyers. 
- Notably, 28 year-olds appear to be less likely to be frequent flyers based on the lower count from the graph. 
- In conclusion, the data suggests that the proportion of frequent and non-frequent flyers varies across different age groups.

```{r}
# grouped bar chart
frequentFlyerStatus <- c("Yes" = "red", "No" = "blue") 
ggplot(data, aes(x = Age, fill = FrequentFlyer)) + 
  geom_bar(position = "dodge") + #Ensures bars for frequent and non-frequent flyers are side by side
  scale_fill_manual(values = frequentFlyerStatus) + #Fills colors for both flyer status values
  labs(title = "Distribution of Frequent Flyer Status by Age", x = "Age", y = "Count") +
  scale_x_continuous(breaks = seq(25, 35, 1))  # Adding ticks for ages 25 to 35
```

# Graph 4 - Sector/Income

-   We can see how the income is distributed by Employment Type (Goverment vs Private).

-   We can see that the income is slightly higher in the private sector.

```{r include=TRUE}
ggplot(data, aes(x = Employment.Type, y = AnnualIncomeUSD, fill = Employment.Type)) +
  geom_boxplot() +
  labs(title = "Box Plot of Annual Income by Employment Type", x = "Employment Type", y = "Annual Income") + # Sets the title and label for x,y axis
    scale_y_continuous(labels = scales::dollar_format(), breaks = seq(from = min(data$AnnualIncomeUSD), to = max(data$AnnualIncomeUSD), by = 2000)) +  # Formats the y-axis labels as dollars and sets breaks
  theme_light() +  
  theme( 
        legend.position = "none",  # Hides the legend 
        plot.title = element_text(hjust = 0.5),  # Center the plot title
       )  

```

# Graph 5 - Age/Income

Age and income is a cool graph to see. We can see how income is distributed by age. - Generally it is believed that income increases with age. - Peak around 40-50 since that is when people are at their peak in their careers.

## What does the data tell us

-   There is not a big difference in the age and the income that these people make.

## By looking at a barplot for each age,

-   There is a slight increase of people who buy insurance as they make more money.
-   It is easier to tell in box plot compared to scatter plot.

```{r fig.width=8, fig.height=4}

p = ggplot(data, aes(x = Age, y = AnnualIncomeUSD)) +
  geom_point(aes(color = as.factor(TravelInsurance))) +  # Convert TravelInsurance to a factor for  color mapping
  scale_x_continuous(breaks = seq(min(data$Age), max(data$Age), by = 1)) +  # Sets the x-axis breaks to increment by 1 (Age) 
  scale_y_continuous(labels = scales::dollar_format(), breaks = seq(from = min(data$AnnualIncomeUSD), to = max(data$AnnualIncomeUSD), by = 2000)) +  # Formats the y-axis labels as dollars and increment by 2k 
  scale_color_manual(values = c("0" = "black", "1" = "orange"), name = "Travel Insurance", labels=c("No ", "Yes")) +  # Sets custom colors for TravelInsurance and customizes legend title
  labs(title = "Age vs Income", x = "Age", y = "Annual Income") +  # Sets the title and label for x,y axis 
  theme_light()
p 
```

## Age/Income as box-plot

```{r  fig.width=8, fig.height=4}
ggplot(data, aes(x = as.factor(Age), y = AnnualIncomeUSD, fill=as.factor(TravelInsurance)  )) + # Plot Age ( min-max) vs AnnualIncomeUSD with fill color as TravelInsurance
  geom_boxplot() +
  labs(title = "Box Plot of Annual Income by Age", x = "Age", y = "Annual Income", fill = "Travel\nInsurance" ) + # Sets the title and label for x,y axis and legend title
  scale_y_continuous(labels = scales::dollar_format(), breaks = seq(from = min(data$AnnualIncomeUSD), to = max(data$AnnualIncomeUSD), by = 2000)) + # Formats the y-axis labels as dollars and increase by 2k 
  # change the text of legend title 
  scale_fill_manual(values = c("0" = "gray", "1" = "green"), labels = c("No", "Yes")) +  # Sets custom colors and labels for the legend
  theme_light() +  
  theme(  plot.title = element_text(hjust = 0.5)  ) # Centers the plot title
```
