---
title: "Project 1"
output:
  powerpoint_presentation: default
  beamer_presentation: default
date: "2024-02-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```




<!-- Load the Dataset
For instance, If you're using a sample dataset like "mtcars," load it using data(mtcars)

Or if you are using a CSV dataset in R, you can use the ‘read.csv’ function, which is part of the base R package.

data <- read.csv("your_file_path.csv")
 

Create at least four different types of ggplot visualizations for the selected variables:
Example, If applicable:

Create a histogram to visualize the distribution of the selected numeric variable. You are encouraged to experiment with bin sizes and labels.
Create a box plot to visualize the median, quartiles, and potential outliers of the selected numeric variable.
Create a density plot to visualize the probability density of the selected numeric variable.
Create a bar plot to visualize the count distribution of the selected categorical variable.
 

Prepare a write-up accompanying your codes, describe the insights you've gained from each plot. -->
 
 
 
# Explain the dataset
The dataset is a travel insurance dataset. It contains information travelers and whether they purchased travel insurance or not and other information (age, income, Employment, etc). 
The dataset contains 10 columns and 1986 rows of information. The columns are:
idx	Age	Employment Type	GraduateOrNot	AnnualIncome	FamilyMembers	ChronicDiseases	FrequentFlyer	EverTravelledAbroad	TravelInsurance

Age- Age Of The Customer
Employment Type- The Sector In Which Customer Is Employed
GraduateOrNot- Whether The Customer Is College Graduate Or Not
AnnualIncome- The Yearly Income Of The Customer In Indian Rupees[Rounded To Nearest 50 Thousand Rupees]
FamilyMembers- Number Of Members In Customer's Family
ChronicDisease- Whether The Customer Suffers From Any Major Disease Or Conditions Like Diabetes/High BP or Asthama,etc.
FrequentFlyer- Derived Data Based On Customer's History Of Booking Air Tickets On Atleast 4 Different Instances In The Last 2 Years[2017-2019].
EverTravelledAbroad- Has The Customer Ever Travelled To A Foreign Country[Not Necessarily Using The Company's Services]
TravelInsurance- Did The Customer Buy Travel Insurance Package During Introductory Offering Held In The Year 2019.


# Load the Dataset 
```{r include=TRUE}
data = read.csv("./TravelInsurancePrediction.csv")

colnames(data)
head(data)

```


# reshaping data 
Data is already in long format. 
Data for income is given in Indian Rupees. We will create a new column to convert it to USD. Column name will be AnnualIncomeUSD
```{r include=TRUE}
RUPEE_TO_USD = 0.012
# add a new column for income in USD
data$AnnualIncomeUSD <- data$AnnualIncome * RUPEE_TO_USD
head(data)

```

# Graph 2 - Age/Income
```{r include=TRUE}
library(ggplot2)
library(reshape2)

ggplot(data, aes(x = Age, y = AnnualIncomeUSD)) +
  geom_point(aes(color = TravelInsurance)) +
  scale_x_continuous(breaks = seq(min(data$Age), max(data$Age), by = 1)) +  # Sets the x-axis breaks to increment by 1
  scale_y_continuous(labels = scales::dollar_format(), breaks = seq(from = min(data$AnnualIncomeUSD), to = max(data$AnnualIncomeUSD), by = 50000)) +  # Formats the y-axis labels as dollars and sets breaks
  labs(title = "Age vs Income",
       x = "Age",
       y = "Annual Income")


ggplot(data, aes(x = as.factor(Age), y = AnnualIncomeUSD)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::dollar_format(), breaks = seq(from = min(data$AnnualIncomeUSD), to = max(data$AnnualIncomeUSD), by = 50000)) +  # Formats the y-axis labels as dollars and sets breaks
  labs(title = "Box Plot of Annual Income by Age",
       x = "Age",
       y = "Annual Income")
```


# Graph 3 - Sector/Income 
```{r include=TRUE}

ggplot(data, aes(x = Employment.Type, y = AnnualIncomeUSD)) +
  geom_boxplot() +
  labs(title = "Box Plot of Annual Income by Employment Type",
       x = "Employment Type",
       y = "Annual Income")



```




 